---
title: "itsjon_OriginalHomeworkCode_04"
author: "Jonathan Zhang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

**Bootstrapping Standard Errors and CIs for Linear Models**

```{r}
# Load data
data <- read.csv("C:/Users/Richard/Desktop/AN588 R Studio/KamilarAndCooperData.csv")

# Create log-transformed variables
data$logHR <- log(data$HomeRange_km2)
data$logBM <- log(data$Body_mass_female_mean)

# Fit linear model
model <- lm(logHR ~ logBM, data = data)

# View coefficients
summary(model)$coefficients
```

*Estimate the standard error for each of your β coefficients as the standard deviation of the sampling distribution from your bootstrap and determine the 95% CI for each of your β coefficients based on the appropriate quantiles from your sampling distribution.*

```{r}
set.seed(123)  # for reproducibility
n_boot <- 1000
boot_coefs <- replicate(n_boot, {
  boot_sample <- data[sample(1:nrow(data), replace = TRUE), ]
  coef(lm(logHR ~ logBM, data = boot_sample))
})

# Transpose for easier handling
boot_coefs <- t(boot_coefs)
colnames(boot_coefs) <- c("Intercept", "Slope")

# Estimate SE and CI
boot_se <- apply(boot_coefs, 2, sd)
boot_ci <- apply(boot_coefs, 2, quantile, probs = c(0.025, 0.975))

# Display results
boot_se
boot_ci
```

```{r}
# Full model SEs
full_se <- summary(model)$coefficients[, "Std. Error"]

# Full model CIs
confint(model)

# Comparison
comparison <- data.frame(
  Coefficient = c("Intercept", "Slope"),
  Full_SE = full_se,
  Bootstrap_SE = boot_se,
  Full_CI_Low = confint(model)[,1],
  Full_CI_High = confint(model)[,2],
  Boot_CI_Low = boot_ci[1, ],
  Boot_CI_High = boot_ci[2, ]
)
comparison
```

**EXTRA CREDIT FUNCTION**

```{r}
bootstrap_lm <- function(d, m, conf.level = 0.95, n = 1000) {
  f <- as.formula(m)
  full_model <- lm(f, data = d)
  full_summary <- summary(full_model)
  full_ci <- confint(full_model, level = conf.level)
  
  coefs <- replicate(n, {
    boot_sample <- d[sample(1:nrow(d), replace = TRUE), ]
    coef(lm(f, data = boot_sample))
  })
  
  coefs <- t(coefs)
  boot_means <- colMeans(coefs)
  boot_se <- apply(coefs, 2, sd)
  alpha <- (1 - conf.level) / 2
  boot_ci <- apply(coefs, 2, quantile, probs = c(alpha, 1 - alpha))
  
  result <- data.frame(
    Coefficient = names(coef(full_model)),
    Full_Estimate = coef(full_model),
    Full_SE = full_summary$coefficients[, "Std. Error"],
    Full_CI_Low = full_ci[, 1],
    Full_CI_High = full_ci[, 2],
    Boot_Mean = boot_means,
    Boot_SE = boot_se,
    Boot_CI_Low = boot_ci[1, ],
    Boot_CI_High = boot_ci[2, ]
  )
  return(result)
}

# Example usage
bootstrap_lm(data, "logHR ~ logBM")
```

**EXTRA EXTRA CREDIT: Graphs Over Increasing Bootstraps**

```{r}
library(ggplot2)
```

```{r}
bootstrap_curve <- function(d, m, max_n = 200, step = 10) {
  f <- as.formula(m)
  full_model <- lm(f, data = d)
  beta_vals <- coef(full_model)
  
  results <- data.frame()
  
  for (n in seq(10, max_n, by = step)) {
    coefs <- replicate(n, {
      boot_sample <- d[sample(1:nrow(d), replace = TRUE), ]
      coef(lm(f, data = boot_sample))
    })
    coefs <- t(coefs)
    ci <- apply(coefs, 2, quantile, probs = c(0.025, 0.975))
    means <- colMeans(coefs)
    
    for (i in 1:length(means)) {
      results <- rbind(results, data.frame(
        Coefficient = names(means)[i],
        N = n,
        Mean = means[i],
        CI_Low = ci[1, i],
        CI_High = ci[2, i],
        Full_Beta = beta_vals[i]
      ))
    }
  }
  return(results)
}

# Run the curve generator
boot_plot_data <- bootstrap_curve(data, "logHR ~ logBM")

# Plot
bootplot_graph <- ggplot(boot_plot_data, aes(x = N)) +
  geom_line(aes(y = Mean, color = Coefficient)) +
  geom_ribbon(aes(ymin = CI_Low, ymax = CI_High, fill = Coefficient), alpha = 0.2) +
  geom_hline(aes(yintercept = Full_Beta, color = Coefficient), linetype = "dashed") +
  labs(title = "Bootstrapped Beta Estimates vs. Number of Resamples",
       y = "Beta Value", x = "Number of Bootstraps") +
  theme_minimal(base_size = 14)
bootplot_graph
```
